---
title: 'Using Linguistic Markers'
author: 'Mark Newman'
date: '2021-04-29'
editor_options: 
  chunk_output_type: console
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Linguistic Markers}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
---

# Background

Linguistic Markers are tokens that are indicative of some category.
When they appear in text, that text is assumed to inherit some of the categorie's meaning.

Take the word "bad" as an example.
This word is may be associated with the category _negative_ tone.
When "bad" appears in the sentence

> It was bad.

The sentence can be assumed to, at least partially, belong to the category _negative_ tone.

Additional features of linguistic markers include:

* Multi-category markers
* Marker groups
* Multiple calculation metrics
* Context defined.

# Basic Walkthrough

Assemble a corpus.

```{r}
corpus <-
  data.frame(
    id = c(1,2,3),
    text = c(
      'It was a bad day, like really bad.',
      'I am glad it is a sunny day.',
      'I am glad it is not a bad day.'))
```

Select your marker file.
For help with your own marker file, see the vignette _Making Marker Files_.

```{r}
path <- system.file("extdata", "sample_markers.csv", package = "LinguisticMarkers")
```

Calculate the linguistic markers

```{r}
library(LinguisticMarkers)

enriched_corpus <- 
  calculate_markers(
    corpus = corpus,
    file = path,
    id = 'id',
    text = 'text',
    show_progress = F)
```

Look at the measures

```{r message = F, warning = F, results = 'hold'}
library(dplyr)
library(knitr)
library(kableExtra)

enriched_corpus %>%
  select(id, WordCount, WordCountUnique, MarkerCount, MarkerCountUnique) %>%
  arrange(id) %>%
  kable(
    caption = 'Basic Stats',
    col.names = c(
      'ID',
      'Count',
      'Unique',
      'Count',
      'Unique')) %>%
  kable_styling() %>%
  add_header_above(c(' ' = 1, 'Word' = 2, 'Marker' = 2))

enriched_corpus %>%
  select(id, negative, negativeUnique, positive, positiveUnique) %>%
  arrange(id) %>%
  kable(
    caption = 'Category Stats',
    col.names = c(
      'ID',
      'Count',
      'Unique',
      'Count',
      'Unique')) %>%
  kable_styling() %>%
  add_header_above(c(' ' = 1, 'Negative' = 2, 'Positive' = 2))
```

Convert the markers into measures.

There are three different ways to convert markers into measures useful in an analysis:

* Ratio
* _Unique_ category count / Total word count.
* Category count / Total word count.

